Case based Arabic Morphological Analysis with Compositional Non-deterministic Automata

### 2010-06-25 tex/sarf.tex
editing

We hypothesise that many NLP case studies need the 
morphological analyser to answer simple queries that do not need 
all this complexity and accuracy and thus most of the work can be ignored. 
For example, if the query is interested in proper names and the 
prefix in question now can only connect to a verb, we can provide an answer
without completing the analysis. 
We find evidence to our hypothesis in~\cite{Maamouri:10} where the 
addition of a new corpus to the Arabic Tree Bank~\cite{Maamouri:04}
with features demanding resolution at an abstraction level
higher than the text itself
led to a refined analyzer~\cite{Kulick:10}.

In this paper, we present Sarf, a {\em novel case-based efficient morphological 
analyzer} that uses parallel compositional non-deterministic Automata 
driven by a case based controller.
Each alive machine in Sarf represents a valid morphological analysis. 
Sarf keeps alive all possible analyses of the text and gives opportunity for the 
controller to intervene at every single input character. 
The decisions of the case-based controller can thus prune false positives early 
in the run. 
Each alive machine in Sarf takes as input one character at a time 
from a text stream. Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
To out knowledge, This solves the ``run-on'' words problem.

@@

\subsection{Contributions}
% does not have to be a subsection, 
%just to highlight it as an item we should not miss

@@

Many morphological analyzers exist~\cite{sWe limit our comparison to related wowour discussion of related work toLots of work

>>>>

We hypothesise that many NLP case studies need the 
morphological analyser to answer simple queries that do not need 
high complexity and accuracy at the low morphological analysis level 
and that can often compensate for tolerable inaccuracy at higher level. 
For example, if the query is interested in proper names and the 
prefix in question in the analysis can only connect to a verb, 
we can provide an answer without completing the analysis. 
%TODO A lso if a word such as \RL{wasiim} was name was ambiguous and we were looking for a sequence of names, then 
We find evidence to our hypothesis in~\cite{Maamouri:10} where the 
addition of a new corpus to the Arabic Tree Bank~\cite{Maamouri:04}
with features demanding resolution at an abstraction level
higher than the text itself
led to a refined analyzer~\cite{Kulick:10}.  
We also find evidence in~\cite{Habash:06} that different types of 
morphological analyses behave differently on the same case study. 
We strongly beleive and we find evidence in our results that a 
case-based controller intervening at every decision is necessary to 
guide, use, prune, and refine the morphological analysis.

\subsection{Sarf}
\label{sec:intro:sarf}

In this paper, we present Sarf, a {\em novel case-based efficient
morphological analyzer} that uses parallel compositional 
non-deterministic automata driven by a case based controller.
Each alive machine in Sarf represents a valid morphological analysis. 
Sarf keeps alive all possible analyses of the text and gives 
opportunity to the controller to intervene at every single input 
character. 
The decisions of the case-based controller can thus prune false 
positives early in the run. 
Each alive machine in Sarf takes as input one character at a time 
from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
To out knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

@@

%\subsection{Contributions}

We make the following key contributions. 
\begin{enumerate}
\item {\bf Case-based analysis:}  We deisgned and implemented
a case-based morphological analyzer where a case-based controller
machine controls and guides the analysis. 
\item {\bf Exhaustive analysis:} Since we have a case-based 
controller that can decide on the fly whether an analysis is 
accepted or not, we can afford to keep all valid analyses modulo
those pruned by the case controller. 
We do that using non-deterministic FSAs. 
\item {\bf Recursive affixes:}
We refine the concatenative analysis of 
Buckwalter~\shortcite{Buckwalter:02}  and use recursive affixes. 
This allowed Sarf to retain better compositional part of speech
tags.
\item {\bf Hadith case study:}
We evaluated Sarf using the hadith case study. We were able to
efficiently extract the sequence of narrators into a complex
data structure with three levels of hierarchy with high accuracy. 
\end{enumerate}

% does not have to be a subsection, 
%just to highlight it as an item we should not miss

@@

Many morphological analyzers exist~\cite{Sughaiyer:04}. 
Some use specialized part of speech tag sets~\cite{Khoja:01,Darwish:02}.

Buckwalter: used lexicon, added proper nouns, word based, recursive affixes

MAGEAD

Beeseley: many compiled FSAs compared to compositions of three FSAs

SAMA: word based, treebank specific,

MADA+TOKAN: no controller, we do not need a tokenizer

AMIRA: machine learning...

ElixirFM:
